---
title: "Logistic regression of TBNI action categories vs seagrass"
format: 
  html:
    code-fold: true
editor: source
lightbox: true

execute: 
  warning: false
  message: false
  echo: true
---

```{r}
#| label: setup
library(tidyverse)
library(patchwork)
library(MASS)
library(marginaleffects)
library(here)

dat <- read.csv(here('data/tbm_combined_catch_env_factors.csv'))

midscr <- 39

cols <- c('#CC3231', '#E9C318', '#2DC938')
labs <- c('On Alert', 'Caution', 'Stay the Course')

tomod <- dat %>% 
  dplyr::select(Reference, month, year, Season, TBEP_seg, 
           FLUCCSCODE, areas, bottom, DominantVeg, bveg, Shore, 
           BvegCovBin, StartDepth, BottomVegCover, BycatchQuantity, 
           TBNI_Score, acres, Non, HA, TH, SAV, 
           Alg, RU, temperature, salinity, dissolvedO2) %>% 
  dplyr::mutate(
    Action = findInterval(TBNI_Score, c(32, 46)),
    outcome = factor(Action, levels = c('0', '1', '2'), labels = cols),
    outcome = as.character(outcome),
    Action = factor(Action, levels = c('0', '1', '2'), labels = labs, ordered = T), 
    TBEP_seg = factor(TBEP_seg, levels = c('OTB', 'HB', 'MTB', 'LTB')),
    grmid = ifelse(TBNI_Score > midscr, 1, 0)
  )
```

This document provides some examples of the likelihood of obtaining the three TBNI action categories (stay the course, caution, on alert) based on seagrass cover.  The intent is to understand potential targets for seagrass cover to obtain a desired TBNI outcome.  Whether it does that is another question.

### Exploratory plots

```{r}
#| fig-height: 7
#| fig-width: 11
#| fig-cap: "TBNI Action categories and scores compared to seagrass cover from the FIM data or seagrass patch size from SWFWMD."
p1 <- ggplot(tomod, aes(x = BottomVegCover, y = TBNI_Score)) + 
  geom_point(aes(color = Action), show.legend = F) + 
  scale_color_manual(values = cols) +
  geom_smooth(method = 'lm', se = F, formula = y ~ x) +
  facet_wrap(~TBEP_seg, ncol = 4) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 8)
  ) +
  labs(
    x = 'Bottom Veg Cover (%)',
    y = 'TBNI Score'
  )

p2 <- ggplot(tomod, aes(x = Action, y = BottomVegCover)) + 
  geom_boxplot(aes(fill = Action), show.legend = F) + 
  scale_fill_manual(values = cols) +
  facet_wrap(~TBEP_seg, ncol = 4) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8)
  ) +
  labs(
    x = 'Action',
    y = 'Bottom Veg Cover (%)'
  )

p3 <- ggplot(tomod, aes(x = acres / 1000, y = TBNI_Score)) + 
  geom_point(aes(color = Action), show.legend = F) + 
  scale_color_manual(values = cols) +
  geom_smooth(method = 'lm', se = F, formula = y ~ x) +
  facet_wrap(~TBEP_seg, scales = 'free_x', ncol = 4) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 8)
  ) +
  labs(
    x = 'Patch Acres (x1000)',
    y = 'TBNI Score'
  )

p4 <- ggplot(tomod, aes(x = Action, y = acres / 1000)) + 
  geom_boxplot(aes(fill = Action), show.legend = F) + 
  scale_fill_manual(values = cols) +
  facet_wrap(~TBEP_seg, scales = 'free_y', ncol = 4) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8)
  ) +
  labs(
    x = 'Action',
    y = 'Patch Acres (x1000)'
  )

p1 + p2 + p3 + p4 + plot_layout(ncol = 2, axis_titles = 'collect')
```

### Binomial logistic regression

```{r}
#| fig-height: 6
#| fig-width: 8
#| fig-cap: !expr paste("Likelihood of obtaining a TBNI score >", midscr, " as the midpoint between all action categories based on seagrass cover from the FIM data or seagrass patch size from SWFWMD") 
mod <- glm(grmid ~ BottomVegCover*TBEP_seg, data = tomod, family = 'binomial')

trgs <- tomod %>% 
  dplyr::select(TBEP_seg, BottomVegCover) %>%
  reframe(
    BottomVegCover = seq(min(BottomVegCover, na.rm = T), max(BottomVegCover, na.rm = T), length.out = 100),
    .by = TBEP_seg
  )
lnprds <- predict.glm(mod, type = 'response', newdata = trgs, se.fit = T)
tolns <- trgs |> 
  mutate(
    prd = lnprds$fit,
    hival = lnprds$fit + 1.96 * lnprds$se.fit,
    loval = lnprds$fit - 1.96 * lnprds$se.fit
  )

p1 <- ggplot(tolns, aes(x = BottomVegCover)) +
  geom_ribbon(aes(ymin = loval, ymax = hival), alpha = 0.2) +
  geom_line(aes(y = prd)) +
  geom_rug(data = tomod[tomod$grmid == 0, ], sides = 'b') +
  geom_rug(data = tomod[tomod$grmid == 1, ], sides = 't') +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank()
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4) +
  labs(
    x = 'Bottom Veg Cover (%)', 
    y = paste('Probability of TBNI Score >', midscr)
  ) 

mod <- glm(grmid ~ acres*TBEP_seg, data = tomod, family = 'binomial')

trgs <- tomod %>% 
  dplyr::select(TBEP_seg, acres) %>%
  reframe(
    acres = seq(min(acres, na.rm = T), max(acres, na.rm = T), length.out = 100),
    .by = TBEP_seg
  )
lnprds <- predict.glm(mod, type = 'response', newdata = trgs, se.fit = T)
tolns <- trgs |> 
  mutate(
    prd = lnprds$fit,
    hival = lnprds$fit + 1.96 * lnprds$se.fit,
    loval = lnprds$fit - 1.96 * lnprds$se.fit
  )

p2 <- ggplot(tolns, aes(x = acres / 1000)) +
  geom_ribbon(aes(ymin = loval, ymax = hival), alpha = 0.2) +
  geom_line(aes(y = prd)) +
  geom_rug(data = tomod[tomod$grmid == 0, ], sides = 'b') +
  geom_rug(data = tomod[tomod$grmid == 1, ], sides = 't') +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank()
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4, scales = 'free_x') +
  labs(
    x = 'Patch Acres (x1000)', 
    y = paste('Probability of TBNI Score >', midscr)
  ) 

p1 + p2 + plot_layout(ncol = 1, axis_titles = 'collect')
```

### Ordinal logistic regression

```{r}
#| fig-height: 7
#| fig-width: 11
#| fig-cap: "Likelihood of obtaining each of three TBNI action categories based on seagrass cover from the FIM data or seagrass patch size from SWFWMD."
mod <- polr(Action ~ BottomVegCover*TBEP_seg, data = tomod, Hess = T)

trgs <- tomod %>% 
  dplyr::select(TBEP_seg, BottomVegCover) %>%
  reframe(
    BottomVegCover = seq(min(BottomVegCover, na.rm = T), max(BottomVegCover, na.rm = T), length.out = 100),
    .by = TBEP_seg
  )

probs <- marginaleffects::predictions(mod, 
                                      newdata = trgs,
                                      type = "probs")
lnprds <- probs %>% 
  dplyr::select(
    Action = group, 
    prd = estimate, 
    loval = conf.low, 
    hival = conf.high, 
    TBEP_seg, 
    BottomVegCover
  ) %>% 
  data.frame()

p1 <- ggplot(lnprds, aes(x = BottomVegCover)) +
  geom_ribbon(aes(ymin = loval, ymax = hival, fill = Action), alpha = 0.5) +
  geom_line(aes(y = prd, color = Action), show.legend = F) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(), 
    legend.position = 'top'
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4, scales = 'free_x') +
  labs(
    x = 'Bottom Veg Cover (%)', 
    y = 'Probability of TBNI Action Category'
  ) 

# area chart
p2 <- ggplot(lnprds, aes(x = BottomVegCover, y = prd, fill = Action)) +
  geom_area(alpha = 0.5) +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(legend.position = 'top') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4, scales = 'free_x') +
  labs(
    x = 'Bottom Veg Cover (%)', 
    y = 'Probability of TBNI Action Category'
  )

mod <- polr(Action ~ acres*TBEP_seg, data = tomod, Hess = T)

trgs <- tomod %>% 
  dplyr::select(TBEP_seg, acres) %>%
  reframe(
    acres = seq(min(acres, na.rm = T), max(acres, na.rm = T), length.out = 100),
    .by = TBEP_seg
  )

probs <- marginaleffects::predictions(mod, 
                                      newdata = trgs,
                                      type = "probs")
lnprds <- probs %>% 
  dplyr::select(
    Action = group, 
    prd = estimate, 
    loval = conf.low, 
    hival = conf.high, 
    TBEP_seg, 
    acres
  ) %>% 
  data.frame()

p3 <- ggplot(lnprds, aes(x = acres / 1000)) +
  geom_ribbon(aes(ymin = loval, ymax = hival, fill = Action), alpha = 0.5) +
  geom_line(aes(y = prd, color = Action), show.legend = F) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(), 
    legend.position = 'top'
  ) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4, scales = 'free_x') +
  labs(
    x = 'Patch Acres (x1000)', 
    y = 'Probability of TBNI Action Category'
  ) 

# area chart
p4 <- ggplot(lnprds, aes(x = acres / 1000, y = prd, fill = Action)) +
  geom_area(alpha = 0.5) +
  scale_fill_manual(values = cols) +
  theme_minimal() +
  theme(legend.position = 'top') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  facet_wrap(~TBEP_seg, ncol = 4, scales = 'free_x') +
  labs(
    x = 'Patch Acres (x1000)', 
    y = 'Probability of TBNI Action Category'
  )

p1 + p2 + p3 + p4 + plot_layout(ncol = 2, guides = 'collect', axis_titles = 'collect') & theme(legend.position = 'top')
```

